---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r, libraries, include=FALSE}

library(tidyverse)
library(clccr)

path_to_folder <- here::here("data-raw/example_data")

```


# clccr

<!-- badges: start -->
<!-- badges: end -->

The clccr package provides a set of functions to compute the commodity life cycle costing (or CLCC) indicator, a measure of resource use by a product/process during its life cycle. The indicator can be thought as a life cycle costing (LCC) indicator developed like it were and life cycle assessment (LCA) indicator: costs are treated like characterization factors, while the classification ad characterization phases are carried out like a conventional Life Cycle Impact Assessment (LCIA). Unlike conventional LCC - which takes into account all costs related to the life cycle of a product - the proposed indicator only considers costs related to natural resource use.

## Installation

You can install the development version of clccr like so:

``` {r install, eval=FALSE}
devtools::install_github("giuliomela/clccr")
```

## Examples

These are basic examples which show you how to use the `clccr` package to compute the CLCC indicator.
The baseline indicator is calculated with the function `clcc`.

```{r example1}
library(clccr)

# computing the CLCC indicator for three example inventories (cars)

clcc(path = path_to_folder)


```

Setting the `critical` argument to `TRUE`, function `clcc` returns the CLCC indicator computed
considering critical materials only. The output tibble provides also the relative share of 
critical materials on the baseline CLCC.

```{r example2}

# computing the critical CLCC indicator

clcc(path = path_to_folder, critical = TRUE)

```

Even though data on reference prices cannot be updated programmatically since not all source have
an API, it is possible to extract a tibble with reference prices and minimum and maximum values (used
in Monte Carlo simulations) of the last 10-year period. The user can  get information on price levels and data sources in the `clccr::clcc_prices_ref` tibble with data sources.

```{r example3}

# A tibble with information on prices used and data sources
prices <- clcc_prices_ref

head(prices, 5)

```

The `clcc` function returns the total CLCC indicator (computed taking into account all material and energy flows) but the user might be interested in having information on the relative importance of each flow on total baseline or critical CLCC. The `clcc_detail` function fills this gap. It returns a tibble with the relative share of each material or energy flow on the CLCC indicator. Such shares can be computed for either the baseline and critical CLCC indicators as well as for the entire life-cycle or just one of the life-cycle phases.

```{r example4}

detail_info <- clcc_detail(path = path_to_folder,
                           critical = FALSE,
                           phase_of_int = "total",
                           collapse_share = 0.9)

head(detail_info, 10)


```

The `clccr` package also allows to run a Monte Carlo simulation, given the high volatility that characterizes
market prices. Minimum and maximum prices are used, for each commodity, to generate random values using 
triangular distributions. Function `clcc_mc` runs the simulations and allows to pick the desired
number of simulations with the `rep` parameter (default is `rep = 10000`) and the life cycle phase to consider
(default is `"total"`). The function returns a tibble nested tibble containing project and object names, baseline clcc, all simulations, and the probability that the indicator is lower than the baseline (probability
computed using the empirical distribution function).

```{r example5}

# Monte Carlo simulation

rep <- 1000 # number of simulations (default is 10,000)
phase <- "total" # the life cycle phase for which running the simulation

clcc_mc(path = path_to_folder, rep = rep)

```

Setting the argument `prob_inf_alt` to `TRUE` (default is `FALSE`), `clcc_mc` returns a tibble  with the probability that each object's CLCC is lower than that of every other object (also across projects, even if it is not always appropriate). Such probabilities are computed, for each object pair, using the empirical distribution function computed on the differences between all the simulated values for each object.  

```{r example6}

# Monte Carlo simulation returning the probability that an object's CLCC is lower/higher than that of the
# other alternatives

clcc_mc(path = path_to_folder, rep = rep, prob_inf_alt = TRUE)

```

Finally, the package allows the user to export reference prices for both the baseline and critical indicators to a .csv file compatible with SimaPro, one of the most popular LCA software. To accomplish this task the user can call the `simapro_export` function. Such function has no arguments and, once called, automatically opens a pop-up windows that asks the user where to save the file.
