---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r, libraries, include=FALSE}

library(tidyverse)
library(clccr)

path_to_folder <- here::here("data-raw/example_data")

path_to_weights <- here::here("data-raw/correzione_coke.xlsx")

```


# clccr

<!-- badges: start -->
<!-- badges: end -->

The clccr package provides a set of functions to compute the commodity life cycle costing (or CLCC) indicator, a measure of resource use by a product/process during its life cycle. The indicator can be thought as a life cycle costing (LCC) indicator developed like it were and life cycle assessment (LCA) indicator: costs are treated like characterization factors, while the classification ad characterization phases are carried out like a conventional Life Cycle Impact Assessment (LCIA). Unlike conventional LCC - which takes into account all costs related to the life cycle of a product - the proposed indicator only considers costs related to natural resource use.

## Installation

You can install the development version of clccr like so:

``` {r install, eval=FALSE}
devtools::install_github("giuliomela/clccr")
```

## Examples

These are basic examples which show you how to use the `clccr` package to compute the CLCC indicator.
The baseline indicator is calculated with the function `clcc`. The function returns a list of two elements. The first, `table`, is a tibble with the calculation result, while the second, `plot`, is a `ggplot` plot object representing the results. The `table` element contains both the baseline and the critical CLCC indicators.

```{r example1}
library(clccr)

# computing the CLCC indicator for three example inventories (cars)

res <- clcc(data_path = path_to_folder
            #path_weights = path_to_weights
            )

res[["table"]]
```

The default plot looks like:

```{r example1-plot}

res[["plot"]]

```


Even though data on reference prices cannot be updated programmatically since not all source have
an API, it is possible to extract a tibble with reference prices and minimum and maximum values (used
in Monte Carlo simulations) of the last 10-year period. The user can  get information on price levels and data sources in the `clccr::clcc_prices_ref` tibble with data sources.

```{r example3}

# A tibble with information on prices used and data sources
prices <- clcc_prices_ref

head(prices, 5)

```

The `clcc` function returns the total CLCC indicator (computed taking into account all material and energy flows) but the user might be interested in having information on the relative importance of each flow on total baseline or critical CLCC. The `clcc_detail` function fills this gap. It returns a list containing a tibble with the relative share of each material or energy flow on the CLCC indicator (`table`) and a `ggplot` object (`plot`). Such shares can be computed for either the baseline and critical CLCC indicators (setting the `detail` parameter to either `TRUE` or `FALSE`) as well as for the entire life-cycle or just one of the life-cycle phases.

```{r example4}

detail_info <- clcc_detail(data_path = path_to_folder,
                           #weights_path = path_to_weights,
                           critical = FALSE,
                           phase_of_int = "total",
                           collapse_share = 0.9)

head(detail_info[["table"]], 10)


```

Also in this case the function returns a default plot of the results.

```{r example4-plot}

detail_info[["plot"]]

```

The package provides two alternative versions of the CLCC indicator, taking into account critical materials only. The first alternative uses the latest version of the European Union list of critical materials, while the second the latest version of the list provided by the International Energy Agency.
Some materials are considered critical by the European Commission of the International Energy Agency (IEA) but are not considered as individual flows by SimaPro. For example the SimaPro flow "Coal, hard" contains "coking coal" which is critical. The same applies for silicon: some of the "Sand" flow is used to produce silicon metal, which is a critical material. For this reason, adjustments are needed. LCA practitioners have to manually calculate the extent to which "coal, hard" and "sand" flows are used for coking coal and silicon metal production respectively. The `clcc`, `clcc_mc` and `clcc_detail` functions all allow the use to provide a table of weights representing the share of "coal, hard" and "sand" flows used to produce coking coal and silicon metal for each object and phase considered. The table must be provided in `.xlsx` format with four columns containing: the name of the object (must be the same used to name the inventories), the name of the substance, the name of the phase (the same used in the inventories) and the weight.

```{r example-weights}

res_weights <- 
  clcc(
    data_path = path_to_folder,
    use_weights = TRUE,
    weights_path = path_to_weights
  )

res_weights[["table"]]

```

The `clccr` package also allows to run a Monte Carlo simulation, given the high volatility that characterizes
market prices. Minimum and maximum prices are used, for each commodity, to generate random values using 
triangular distributions. Function `clcc_mc` runs the simulations and allows to pick the desired
number of simulations with the `rep` parameter (default is `rep = 10000`) and the life cycle phase to consider
(default is `"total"`). The function returns a tibble nested tibble containing project and object names, baseline clcc, all simulations, and the probability that the indicator is lower than the baseline (probability computed using the empirical distribution function). Similarly to the `clcc()`, also `clcc_mc()` returns a list: a tibble with the results and a plot displaying the simulation results. Please refer to the `clcc_mc()` documentation to see which are the plotting alternatives available.

```{r example5}

# Monte Carlo simulation

rep <- 1000 # number of simulations (default is 10,000)
phase <- "total" # the life cycle phase for which running the simulation

mc_res <- clcc_mc(data_path = path_to_folder, 
                  #weights_path = path_to_weights,
                  rep = rep)

mc_res[["table"]]

```

Setting the argument `prob_inf_alt` to `TRUE` (default is `FALSE`), `clcc_mc` returns a tibble  with the probability that each object's CLCC is lower than that of every other object (also across projects, even if it is not always appropriate). Such probabilities are computed, for each object pair, using the empirical distribution function computed on the differences between all the simulated values for each object.  

```{r example6}

# Monte Carlo simulation returning the probability that an object's CLCC is lower/higher than that of the
# other alternatives

mc_res2 <- clcc_mc(data_path = path_to_folder, 
                   #path_weights = path_to_weights,
                   rep = rep, prob_inf_alt = TRUE)

mc_res2[["table"]]

```

The package also allows the user to export reference prices for both the baseline and critical indicators to a .csv file compatible with SimaPro, one of the most popular LCA software. To accomplish this task the user can call the `simapro_export()` function. Such function has no arguments and, once called, automatically opens a pop-up windows that asks the user where to save the file.

Finally, function `copy_xlsx_files()` allows the user to copy raw SimaPro LCA inventories from a folder to another and, at the same time, changing the file extension from `.XLSX` to `.xlsx` if needed. SimaPro-generated spreadsheets often have this problem, which can prevent the inventories from being loaded correctly by the `clcc()` or the `clcc_mc()` functions.
