---
title: "clccr-vignette"
author: Giulio Mela
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{clccr-vignette}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup, include=FALSE}
library(clccr)
```

# Introduction

The `clccr` package provides a set of functions to compute the *commodity life cycle costing* (or CLCC) indicator, a measure of resource use by a product/process during its life cycle. The indicator can be thought as a life cycle costing (LCC) indicator developed like it were and life cycle assessment (LCA) indicator: costs are treated like characterization factors, while the classification ad characterization phases are carried out like a conventional Life Cycle Impact Assessment (LCIA). Unlike conventional LCC - which takes into account all costs related to the life cycle of a product - the proposed indicator only considers costs related to natural resource use.

In other words, the CLCC indicator quantifies, in monetary units, the extent to which a product exploits natural resources during its life cycle and can represent an important tool for policy making, highlighting - together with conventional LCA - non-renewable resource consumption by products.

The CLCC indicator is based on market prices (or their proxies, as detailed below), which are a measure of resource scarcity, and can be interpreted as a measure of the extent to which a product/process exploits natural resources during its life cycle. Other LCA indicators of resource use/depletion are available in the literature but only a few of them are of economic type. Furthermore, most of them rely on strong assumptions regarding available reserves, current and future rates of consumption and supply and demand, undermining the robustness of their results. Our proposed indicator avoids this problem using market prices as a proxy for resource scarcity. Furthermore, market prices include information on both demand and supply sides and, at least partially, expectation on their future trends, at least in the medium run.

The starting point for the calculation of the indicator is the list of all raw materials and energy flows considered in the LCA indicator Mineral and Fossil and Renewable Resource Depletion from the International Reference Life Cycle Data System (ILCD), while the second step is the identification, for every material and energy flow, of the corresponding market price[^1].

[^1]: The list of raw materials considered can be expanded if needed, even though, at the moment, this operation cannot be done dynamically using `clccr` package functions.

The package allows also the calculation of an alternative version of the indicator, based on the materials included in the European Commission's list of critical raw materials. Such materials are very important for technology-intensive applications but are - at the same time - characterized by high supply risks.

All price data come from open-source databases such as the [International Monetary Fund's Primary Commodity Prices database](https://www.imf.org/en/Research/commodity-prices) or [Eurostat](https://ec.europa.eu/eurostat/data/database). Nonetheless, it is not possible to identify a reference market prices for most of the raw materials considered: for this reason, for such materials, market price is approximated with export unit values at world level. Trade data come from the United Nations' [Comtrade](https://comtrade.un.org/), European Union's [Comext](https://ec.europa.eu/eurostat/web/international-trade-in-goods/data/focus-on-comext) and the United States' [USITC](https://dataweb.usitc.gov/) databases. More specifically, most of unit values are calculated using four or six-digit trade flow data from Comtrade, except for rare earths materials which require a higher level of detail, which is provided by Comext and USITC databases.

Reference prices used for the calculation of the CLCC indicator are, for each commodity, the average of price of the previous ten-year period, converted in euros using market exchange rates and adjusted to the price level of the reference year (which can be defined by the user) with the Euro Area GDP deflator. Minimum and maximum prices recorded in the ten-year period considered, are used as a range for the uncertainty assessment through Monte Carlo simulations.

More information on the methodology developed for the calculation of the CLCC indicator is available in @mela2021, as well as an extensive literature review on other measures of resource scarcity in the LCA framework, a practical application of the CLCC indicator, and a comparison between the CLCC indicator with other popular measures of resource scarcity in the life cycle.

# Package installation and load {#installation}

The package can be installed trough the GitHub website of the author:

```{r installation, eval=FALSE}

# installing the package

devtools::install_github("giuliomela/clccr")

# loading the package

library(clccr)
```

# Usage {#usage}

The package provides a function, `clcc` to load SimaPro inventories in .xlsx format, from a folder designated by the user, and to compute the CLCC indicator. Each spreadsheet contains inventory data for every phase of a given product/process. The `clcc` function computes the CLCC indicator for each product and phase using the latest version of reference prices available. Price are updated manually[^2] on a yearly basis by the package developers even though, in the future, a new function allowing the user to directly download the series or to change the time period along which average prices are computed.

[^2]: Some price series cannot be accessed through API services and must be downloaded manually.

Below is an example of use of the `clcc` function. The name of each object for which the CLCC indicator is computed is extrapolated from the name of each inventory file.

```{r example-clcc}

library(clccr) # loading the package

# identifying the path the the folder with inventory data

path_to_folder <- here::here("data-raw/example_data") 

clcc_results <- clcc(path = path_to_folder)

clcc_results


```

The `clcc` function can also return a version of the CLCC indicator which takes into account critical materials only, as shown in the next example.

```{r example-clcc-critical}

clcc_critical <- clcc(path = path_to_folder,
                      critical = TRUE)

clcc_critical

```

Even though reference prices are loaded manually into the package folders, it is possible to retrieve the price matrix accessing the database `clcc_prices_ref` as follows

```{r example-price-ref, echo=TRUE}

clcc_prices_ref

```

The `clcc` function returns the total CLCC indicator (computed taking into account all material and energy flows) but the user might be interested in having information on the relative importance of each flow on total baseline or critical CLCC. The `clcc_detail` function fills this gap. It returns a tibble with the relative share of each material or energy flow on the CLCC indicator. Such shares can be computed for either the baseline and critical CLCC indicators as well as for the entire life-cycle or just one of the life-cycle phases. The `phase_of_int` parameter allows the user to select one peculiar phase of the life-cycle, while the `collapse_share` parameter let the use determine the cut-off point to limit the number of commodities returned as output. For example, if `collapse_share` is set to `0.9`, only commoties which share on the CLCC indicator of interest sum up to 90% are reterned and the others are aggregated under the `others` label.

```{r example-clcc-detail}

detail_info <- clcc_detail(path = path_to_folder,
                           critical = FALSE,
                           phase_of_int = "total",
                           collapse_share = 0.9)

head(detail_info, 10)


```



Finally, the `clccr` package provides a function which allows to perform a Monte Carlo analysis to take into account price volatility: `clcc_mc`. The analysis is carried out assuming that reference prices follow a triangular distribution, since no information are available on the true shape of price distributions (even though, in the future, it might be possible to use *distribution fitting* techniques to identify the probability distribution that better describes each price series). The user can specify the number of iterations (through the `rep` parameter, default is 10,000), the life cycle phase to consider and the type of output.

```{r example-mc}

# identifying the path the the folder with inventory data

path_to_folder <- here::here("data-raw/example_data") 

monte_carlo <- clcc_mc(path = path_to_folder,
                       rep = 10000,
                       phase = "total",
                       prob_inf_alt = FALSE)

monte_carlo

```

The standard output of the `clcc_mc` function is a tibble containing project and object names, baseline clcc, all simulations, and the probability that the indicator is lower than the baseline value. If the `prob_inf_alt` parameter is set to `TRUE`, the function also computes the probabilities that each object's CLCC is lower than that of every other object (also across projects, even if it is not always appropriate). Probabilities are always calculated using the empirical distribution function.

```{r example-mc-alt}

monte_carlo_alt <- clcc_mc(path = path_to_folder,
                       rep = 10000,
                       phase = "total",
                       prob_inf_alt = TRUE)

monte_carlo_alt

```

Other functions included in the package, such as `inventory_load_fn` and `start_data_finder_fn` are used within `clcc` and `clcc_mc` and are not meant for independent use.
